---
title: "TOM-I_Analysis EXP1&2 Mixed Effects Logistic Regression"
author: "Elif Poyraz"
date: '2025-02-06'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

Reading the data file from Excel:
```{r echo=FALSE}
library(readxl) #installing the library for reading excel files
library(aod)
library(tidyr)
library(lme4)
library(ggplot2)
library(MASS)
library(performance)
```


```{r}
myCols <- as.character(read_excel("TOM-I_Exp1-2Combined.xlsx",skip=0, n_max = 1, col_names = FALSE)); #defining col names for the database

tomI_Exp1 <- read_excel("TOM-I_Exp1-2Combined.xlsx", skip = 1, col_names = myCols, col_types = c("text", "text", "numeric", "numeric", "date", "date", "numeric", "numeric","numeric", "numeric", "numeric", "numeric","numeric", "numeric", "numeric","numeric"));

TomIMixed_1 <- pivot_longer(tomI_Exp1, cols = c(FB, LD_FB, KNOW, DES), names_to = "Task", values_to = "Dependent")
TomIMixed_2 <- pivot_longer(tomI_Exp1, cols = c(FB, LD_FB, Set_FB, KNOW, DES), names_to = "Task", values_to = "Dependent")#change columns to rows ## added by Yanyi Chen
TomIMixed_3 <- pivot_longer(tomI_Exp1, cols = c(FB, LD_FB), names_to = "Task", values_to = "Dependent")

```


```{r}
tomI_Exp1$AgeGroup <- ifelse(tomI_Exp1$Age >= 60, "5",
                               ifelse(tomI_Exp1$Age >= 48, "4", "3"));

TomIMixed_1$AgeGroup <- ifelse(TomIMixed_1$Age >= 60, "5",
                               ifelse(TomIMixed_1$Age >= 48, "4", "3"));

TomIMixed_2$AgeGroup <- ifelse(TomIMixed_2$Age >= 60, "5",
                               ifelse(TomIMixed_2$Age >= 48, "4", "3"));
```
Summary info on all variables:

```{r echo=FALSE}
summary(tomI_Exp1);
```


Correlations between FB measures
```{r}
cor.test(tomI_Exp1$FB, tomI_Exp1$LD_FB, method="pearson")
```

Correlation Matrix
```{r}
library("Hmisc")
data_sub <- subset(tomI_Exp1, select = -c(SUB_ID, Gender, DOT, DOB))
corr1 = rcorr(as.matrix(data_sub))
corr1

```

```{r}
library(corrplot)
#corrplot(corr1$r, type = 'lower', order = 'hclust', tl.col = 'black',
       #  cl.ratio = 0.2, tl.srt = 45, col = COL2('PuOr', 10))
```

Full model with two interaction terms. Only LD and HD FB. Age categorical.
```{r}

#TomIMixed_1$AgeGroup <- ifelse(TomIMixed_1$Age >= 60, "5",
                               #ifelse(TomIMixed_1$Age >= 48, "4", "3"));

#TomIMixed_2$AgeGroup <- ifelse(TomIMixed_2$Age >= 60, "5",
                               #ifelse(TomIMixed_2$Age >= 48, "4", "3"));

TomIMixed_1_upd <- subset(TomIMixed_1, TomIMixed_1$Task == "FB" | TomIMixed_1$Task == "LD_FB");

TomIMixed_1_upd <- within(TomIMixed_1_upd, {
  SUB_ID <- factor(SUB_ID)
  Task <- factor(Task)
  AgeGroup <- factor(AgeGroup)
});

# estimate the model and store results in m
m <- glmer(Dependent ~  Task + Opp_1 + AgeGroup + Task:Opp_1 + Task:AgeGroup + (1 | SUB_ID), data = TomIMixed_1_upd, family = binomial, control = glmerControl(optimizer = "bobyqa"), nAGQ = 10);

# print the mod results without correlations among fixed effects
print(m, corr = FALSE)

summary(m)

se <- sqrt(diag(vcov(m)))
# table of estimates with 95% CI
(tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *
    se))

# odds ratio?
exp(tab)

```

```{r}

TomIMixed_1_upd2 <- subset(TomIMixed_1, TomIMixed_1$Task == "FB" | TomIMixed_1$Task == "LD_FB" | TomIMixed_1$Task == "KNOW");
TomIMixed_2_upd <- subset(TomIMixed_2, TomIMixed_2$Task != "DES"); # has Setoh task


TomIMixed_1_upd2 <- within(TomIMixed_1_upd2, {
  SUB_ID <- factor(SUB_ID)
  Task <- factor(Task)
  AgeGroup <- factor(AgeGroup)
});

TomIMixed_2_upd <- within(TomIMixed_2_upd, {
  SUB_ID <- factor(SUB_ID)
  Task <- factor(Task)
  AgeGroup <- factor(AgeGroup)
});
```

Full model with two interaction terms. KNOW, LD and HD FB. Age categorical.
```{r age categorical}
# estimate the model and store results in m
m2 <- glmer(Dependent ~ Task + Opp_1 + AgeGroup + Task:Opp_1 + Task:AgeGroup + (1 | SUB_ID), data = TomIMixed_1_upd2, family = binomial, control = glmerControl(optimizer = "bobyqa"), nAGQ = 10);

# print the mod results without correlations among fixed effects
print(m2, corr = FALSE)

summary(m2)

se2 <- sqrt(diag(vcov(m2)))
# table of estimates with 95% CI
(tab2 <- cbind(Est = fixef(m2), LL = fixef(m2) - 1.96 * se2, UL = fixef(m2) + 1.96 *
    se2))

exp(tab2)

```

Full model with two interaction terms. Only LD and HD FB. Age continuous.
Seems continuous model has a better fit for these exact variables. AIC is 5 points lower.
```{r age cont}

# estimate the model and store results in m
m3 <- glmer(Dependent ~ Task + Opp_1 + Age  + Task:Opp_1 +(1 | SUB_ID), data = TomIMixed_1_upd, family = binomial, control = glmerControl(optimizer = "bobyqa"), nAGQ = 10);

# print the mod results without correlations among fixed effects
print(m3, corr = FALSE)

summary(m3)

se3 <- sqrt(diag(vcov(m3)))
# table of estimates with 95% CI
(tab3 <- cbind(Est = fixef(m3), LL = fixef(m3) - 1.96 * se3, UL = fixef(m3) + 1.96 *
    se3))

exp(tab3)

```
```{r age cont}

# estimate the model and store results in m
m5 <- glmer(Dependent ~ Task + Opp_1 + Age  + Task:Opp_1 + Age:Opp_1 +(1 | SUB_ID), data = TomIMixed_1_upd, family = binomial, control = glmerControl(optimizer = "bobyqa"), nAGQ = 10);

# print the mod results without correlations among fixed effects
print(m5, corr = FALSE)

summary(m5)

se5 <- sqrt(diag(vcov(m5)))
# table of estimates with 95% CI
(tab5 <- cbind(Est = fixef(m5), LL = fixef(m5) - 1.96 * se5, UL = fixef(m5) + 1.96 *
    se5))

exp(tab5)

```


Full model with two interaction terms. KNOW, Emma, LD and HD FB. Age continuous.

```{r age continuous}

# estimate the model and store results in m
m4 <- glmer(Dependent ~ Task + Opp_1 + Age + Task:Opp_1 + Task:Age + (1 | SUB_ID), data = TomIMixed_2_upd, family = binomial, control = glmerControl(optimizer = "bobyqa"), nAGQ = 10);

# print the mod results without correlations among fixed effects
print(m4, corr = FALSE)

summary(m4)

se4 <- sqrt(diag(vcov(m4)))
# table of estimates with 95% CI
(tab4 <- cbind(Est = fixef(m4), LL = fixef(m4) - 1.96 * se4, UL = fixef(m4) + 1.96 *
    se4))

exp(tab4)

```

```{r}
myCols1 <- as.character(read_excel("Exp1-2_colnames.xlsx",skip=0, n_max = 1, col_names = FALSE)); #defining col names for the database

tomI_Exp12 <- read_excel("TOM-I_Exp1-2Combined.xlsx", skip = 1, col_names = myCols1, col_types = c("text", "text", "numeric", "numeric", "date", "date", "numeric", "numeric","numeric", "numeric", "numeric", "numeric","numeric", "numeric", "numeric","numeric"));

TomIMixed_12 <- pivot_longer(tomI_Exp12, cols = c(Desire, Knowledge,'Low Demand FB', 'High Demand FB'), names_to = "Task", values_to = "Dependent")
TomIMixed_22 <- pivot_longer(tomI_Exp12, cols = c(Desire, Knowledge, 'Emma Task', 'Low Demand FB', 'High Demand FB'), names_to = "Task", values_to = "Dependent")#change columns to rows ## added by Yanyi Chen

```

```{r}

TomIMixed_12$Task <- factor(TomIMixed_12$Task, levels = c('Desire', 'Knowledge', 'Emma Task', 'Low Demand FB', 'High Demand FB'));

ggplot(data= TomIMixed_12, aes(Opposites, Dependent, color=Task)) +
  geom_point(shape = "|", size = 6, na.rm = TRUE, aes(color = factor(Task))) +
  geom_smooth(method = glm, method.args = list(family = binomial), na.rm = TRUE, 
              formula = y ~ x) +
  coord_cartesian(ylim = c(0, 1), expand = 0) +
  labs(x = "Inhibition", y = "Accuracy") +
  theme_minimal(base_size = 16) +
  theme(axis.line = element_line(color = "gray"),
        axis.ticks = element_line(color = "gray"),
        axis.ticks.length = unit(3, "mm"),
        legend.position = "right")+
  ggtitle("Theory of Mind & Inhibition", subtitle = "The Opposites Task")+
  scale_y_continuous(breaks=seq(0,1,.1))+
  scale_color_manual("Task", values = c("Desire" = "lightsalmon1", "Knowledge" = "brown2","Low Demand FB" = "royalblue", "High Demand FB" = "royalblue4"))

```


```{r}

TomIMixed_22$Task <- factor(TomIMixed_22$Task, levels = c('Desire', 'Knowledge', 'Emma Task', 'Low Demand FB', 'High Demand FB'));

ggplot(data= TomIMixed_22, aes(Opposites, Dependent, color=Task)) +
  geom_point(shape = "|", size = 6, na.rm = TRUE, aes(color = factor(Task))) +
  geom_smooth(method = glm, method.args = list(family = binomial), na.rm = TRUE, 
              formula = y ~ x) +
  coord_cartesian(ylim = c(0, 1), expand = 0) +
  labs(x = "Inhibition", y = "Accuracy") +
  theme_minimal(base_size = 16) +
  theme(axis.line = element_line(color = "gray"),
        axis.ticks = element_line(color = "gray"),
        axis.ticks.length = unit(3, "mm"),
        legend.position = "right")+
  ggtitle("Theory of Mind & Inhibition", subtitle = "The Opposites Task")+
  scale_y_continuous(breaks=seq(0,1,.1))+
  scale_color_manual("Task", values = c("Desire" = "lightsalmon1", "Knowledge" = "brown2","Emma Task"= "skyblue",  "Low Demand FB" = "royalblue", "High Demand FB" = "royalblue4"))

```


```{r}

ggplot(data= TomIMixed_22, aes(Age, Dependent, color=Task)) +
  geom_point(shape = "|", size = 6, na.rm = TRUE, aes(color = factor(Task))) +
  geom_smooth(method = glm, method.args = list(family = binomial), na.rm = TRUE, 
              formula = y ~ x) +
  coord_cartesian(ylim = c(0, 1), expand = 0) +
  labs(x = "Age", y = "Accuracy") +
  theme_minimal(base_size = 16) +
  theme(axis.line = element_line(color = "gray"),
        axis.ticks = element_line(color = "gray"),
        axis.ticks.length = unit(3, "mm"),
        legend.position = "right")+
  scale_color_manual("Task", values = c("Desire" = "lightsalmon1", "Knowledge" = "brown2","Emma Task"= "skyblue",  "Low Demand FB" = "royalblue", "High Demand FB" = "royalblue4"))

```

```{r}
check_model(m3)

```